---
title: "Bayes KCNQ1 Data Process"
author: "Matthew O'Neill and Brett Kroncke"
date: "6/7/2022"
output: 
   html_document:
     toc: true
     toc_depth: 4
     toc_float: false
        #collapse: true
     smooth_scroll: true
     code_folding: hide
     highlight: zenburn #textmate?
     theme: flatly
     #number_sections: true
editor_options: 
   chunk_output_type: console
---


```{r preamble, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# When working in the Bayes R Project, adjust knitting setting from Document -> Project directory to knit without specificying absolute file paths 

source('RMD and Scripts/func_dist_seq.R')

library("nnet")
library("DBI")
library("RSQLite")
library(dplyr)
library(ggplot2)
library(caret)
library(plotrix)
library(glmnet)
library(meta)
library(reshape2)
library(psych)
require(Hmisc)
library(tableone)
library(wCorr)
library(rms)
library(boot)
library(leaps)
library(car)
library(reticulate)
library(rootSolve)
library(pROC)
library(MALDIquant)
library(tidyverse)      
library(lubridate)      
library(fpp2)         
library(zoo)
library(latex2exp)
library(forestplot)
library(ggplot2)

```

# Introduction
This document describes an estimate of LQT1 penetrance from *KCNQ1* missense variants. We incorporate structure, function, and *in silico* predictive models to inform Bayesian penetrance priors. We use this data as input to develop a model of KCNQ1 variant penetrance, and to then investigate patterns of penetrance among four genetic arrhythmias. 

# Data Preparation
Heterozygote data were curated from the literature, retrieved from population databases, or supplied by clinical collaborators.
Data are read in, cleaned, and predictive features are prepared for analysis.
Data are then saved to the `RData` format for use in the Evaluation and Application files.  


```{r data prep, eval = TRUE}

# Load curated literature data as described in the methods 

load('Heterozygote Data/LitFinal_v7.RData')

d <- lit_final

missense <- c("S", "T", "V", "D", "P", "E", "G",  "Y", "R", "W", 
              "F", "H", "N", "A", "K", "Q", "L", "I", "C", "M", 
              "del", "LL", "dup", "MM", "Del", "Ins", "ins")

nonsense <- c("X", "RX","PX", "WX", "TX", "CX", "GX", "VX", 
              "QX", "AX", "EX", "KX", "IX", "fsX", "insGAG", 
              "sp", "", "fs", "splice", "LX", "-")

d$mut_type <- "nonsense"
d[d$mutAA %in% missense, "mut_type"] <- "missense"
lit.nonoverlap.data <- d[,c("var", "resnum", "natAA", "mutAA", "lqt1", "total_carriers", "mut_type")] 
lit.nonoverlap.data$unaff <- lit.nonoverlap.data$total_carriers - lit.nonoverlap.data$lqt1
lit.nonoverlap.data[as.character(lit.nonoverlap.data$mutAA) == as.character(lit.nonoverlap.data$natAA),"mut_type"] <- "synonymous"


```

# Read in Cohort Data
We process the data supplied by tertiary care centers and combine them into a single dataframe. 

```{r Cohort Data, echo = T, results = 'hide'}


# Read in data from each site - manual review removes potential overlap from previously published probands

japan.data <- read.csv('Heterozygote Data/Japan_cohort-v1.csv')

load('Heterozygote Data/KCNQ1_Italy_v3.RData')
italy.data <- italy_final

load('Heterozygote Data/KCNQ1_Paris_v2.RData')
paris.data <- paris_final

# Combine all individuals from cohort by missense variants

cohort.data <- as.data.frame(unique(c(as.character(paris.data$var), as.character(italy.data$var), as.character(japan.data$var))))
names(cohort.data) <- "var"
lit.cohort.data <- as.data.frame(unique(c(as.character(d$var), as.character(cohort.data$var))))
names(lit.cohort.data) <- "var" # lit.cohort.data has the greatest number of variants. 
mut_type <- as.data.frame(c(as.character(d$mutAA), as.character(paris.data$mutAA), as.character(italy.data$mutAA), as.character(japan.data$mutAA)))
native <- as.data.frame(c(as.character(d$natAA), as.character(paris.data$natAA), as.character(italy.data$natAA), as.character(japan.data$natAA)))
resnum <- as.data.frame(c(as.character(d$resnum), as.character(paris.data$resnum), as.character(italy.data$resnum), as.character(japan.data$resnum)))
variants <- as.data.frame(c(as.character(d$var), as.character(paris.data$var), as.character(italy.data$var), as.character(japan.data$var)))
names(variants) <- "var"
names(mut_type) <- "mutAA"
names(native) <- "natAA"
names(resnum) <- "resnum"
mut_type$mut_type <- "nonsense"
mut_type[mut_type$mutAA %in% missense, "mut_type"] <- "missense"
mut_type[,c("var","natAA","resnum")] <- c(as.character(variants$var), as.character(native$natAA), as.character(resnum$resnum))
mut_type$resnum <- as.numeric(mut_type$resnum)
mut_type[mut_type$mutAA == mut_type$natAA & !is.na(mut_type$mutAA) & !is.na(mut_type$natAA), "mut_type"] <- "synonymous"
mut_type <- unique(mut_type)


# Combine all affected and unaffected individuals who carry the same variant

for (variant in cohort.data$var){
  lqt1_I <- 0
  unaff_I <- 0
  lqt1_P <- 0
  unaff_P <- 0 
  lqt1_J <- 0
  unaff_J <- 0
  if (!is.na(match(variant,italy.data$var))) {
    print(variant)
    lqt1_I<-italy.data[italy.data$var == variant,"lqt1_I"]
    print(lqt1_I)
    unaff_I<-italy.data[italy.data$var == variant,"total_carriers_I"] - italy.data[italy.data$var == variant,"lqt1_I"]
  }
  if (!is.na(match(variant,paris.data$var))) {
    print(variant)
    lqt1_P<-paris.data[paris.data$var == variant,"lqt1_P"]
    print(lqt1_P)
    unaff_P<-paris.data[paris.data$var == variant,"total_carriers_P"] - paris.data[paris.data$var == variant,"lqt1_P"]
  }
  if (!is.na(match(variant,japan.data$var))) {
    print(variant)
    lqt1_J<-japan.data[japan.data$var == variant,"lqt1_J"]
    print(lqt1_J)
    unaff_J<-japan.data[japan.data$var == variant,"total_carriers_J"] - japan.data[japan.data$var == variant,"lqt1_J"]
  }
  lqt1 <- lqt1_I + lqt1_P + lqt1_J
  unaff <- unaff_I + unaff_P + unaff_J
  cohort.data[cohort.data$var == variant, "lqt1"] <- lqt1
  cohort.data[cohort.data$var == variant, "unaff"] <- unaff
  cohort.data[cohort.data$var == variant, "total_carriers"] <- unaff + lqt1
}


```

# Combine individuals from cohort and the literature
We next combine clinical cohort observations with those from the literature curation. Potential overlap heterozygotes between the tertiary care centers and publications were removed during the curation to avoid double counting. 

```{r Combine Data, echo = T, results = 'hide'}

for (variant in lit.cohort.data$var){
  lqt1_c <- 0 
  lqt1_l <- 0 
  unaff_c <- 0 
  unaff_l <- 0 
  if (!is.na(match(variant,d$var))) {
    print(variant)
    lqt1_l <- d[d$var == variant,"lqt1"]
    print(lqt1_l)
    unaff_l <- d[d$var == variant,"total_carriers"] - d[d$var == variant,"lqt1"]
  } 
  if (!is.na(match(variant,cohort.data$var))) {
    print(variant)
    lqt1_c <- cohort.data[cohort.data$var == variant,"lqt1"]
    print(lqt1_l)
    unaff_c <- cohort.data[cohort.data$var == variant,"unaff"]
  }
  lqt1 <- lqt1_l + lqt1_c
  unaff<- unaff_l + unaff_c
  lit.cohort.data[lit.cohort.data$var == variant,"lqt1"] <- lqt1
  lit.cohort.data[lit.cohort.data$var == variant,"unaff"] <- unaff
  lit.cohort.data[lit.cohort.data$var == variant,"total_carriers"] <- unaff + lqt1
}

# Merge dataframes 

cohort.data <- unique(merge(cohort.data,mut_type,all = TRUE))
lit.cohort.data <- unique(merge(lit.cohort.data,mut_type,all = TRUE))

# Remove variants with 0 carriers

cohort.data <- cohort.data[cohort.data$total_carriers>0 & !is.na(cohort.data$total_carriers),]
lit.cohort.data <- lit.cohort.data[lit.cohort.data$total_carriers>0 & !is.na(lit.cohort.data$mut_type),]

```


# Assembly of in silico covariates and functional data
Here we merge in silico and functional data with clinical observations for each variant. We will combine these data using an expectation-maximization framework in the subsequent scripts. 

```{r in silico and functional}

# Load in silico covariates

AAs<-c("A","G","I","L","V","M","F","W","P","C","S","T","Y","N","Q","H","K","R","D","E","X")
AA2<-c("A","G","I","L","V","M","F","W","P","C","S","T","Y","N","Q","H","K","R","D","E")

#Cardiacboost data

cb<-read.table('Covariates/cardiacboost_arm_all_possible_mutations-adj.csv', sep = "\t", header = TRUE)
cb<-cb[cb$gene=="KCNQ1",]
cb$var<-as.character(paste(cb$var))
cb$mut_type<-"missense"
cb<-cb[,c("pathogenicity","var","mut_type")] 
names(cb)[1]<-"cardiacboost"

# PSSM data 

pssm<-read.table('Covariates/EMBL_EBI_KCNQ1_PSSM.txt', sep = "", header = TRUE)
pssm_clean <- pssm[,c(1:22)]
pssm_melt<-melt(data = pssm_clean, id.vars = c("resnum", "wtAA"), measure.vars = AA2, variable.name = "mutAA", value.name = "blast_pssm")

# PolyPhen2 Data

pph2<-read.csv('Covariates/pph2-full_kcnq1.txt', sep = '\t', strip.white = TRUE)
pph2_kcnq1_adj <- pph2[,c("pos", "aa1", "aa2", "pph2_class", "pph2_prob", "pph2_FPR", "pph2_TPR", "pph2_FDR")]
names(pph2_kcnq1_adj)[names(pph2_kcnq1_adj) == "pos"] <- "resnum"
names(pph2_kcnq1_adj)[names(pph2_kcnq1_adj) == "aa1"] <- "natAA"
names(pph2_kcnq1_adj)[names(pph2_kcnq1_adj) == "aa2"] <- "mutAA"

# Provean

prov_kcnq1 <- read.table('Covariates/Provean_and_SIFT_KCNQ1.tsv', sep = '\t', header = FALSE)
prov_kcnq1$var <- paste(prov_kcnq1$V5, prov_kcnq1$V4, prov_kcnq1$V6, sep = '')
names(prov_kcnq1) <- c("number", "ENSP1", "ENSP2", "Residue", "native", "mutant", "score", "class", "random", "random1", "random2", "consequence", "score1", "score2", "var")
prov_kcnq1_format <- prov_kcnq1[,c("var", "score")]
names(prov_kcnq1_format)[names(prov_kcnq1_format) == "score"] <- "provean_score"

# Revel

load('Covariates/revel.RData')
revel <- revel_kcnq1
names(revel) <- c("var","revel_score")


# Combine in silico predictors into a single dataframe

in.silico <- merge(pph2_kcnq1_adj, pssm_melt, all = TRUE)
in.silico <- unique(in.silico)
in.silico <- in.silico[!is.na(in.silico$pph2_class),] # removes synonymous variant from list 

in.silico$var <- paste(in.silico$natAA, in.silico$resnum, in.silico$mutAA, sep = "")
in.silico <- merge(cb, in.silico, all = TRUE)
in.silico <- unique(in.silico)
in.silico <- merge(in.silico, prov_kcnq1_format, all = TRUE)
in.silico <- unique(in.silico)
in.silico <- merge(in.silico, revel, all = TRUE)
in.silico <- unique(in.silico)
in.silico <- in.silico[,c("var", "resnum", "pph2_prob","blast_pssm", "provean_score", "revel_score", "cardiacboost", "mut_type")]

# Only take one entry per variant

in.silico <- distinct(in.silico, var, .keep_all = TRUE)


# Curated Functional Data from Syncropatch and Manual Patchclamp data 

q1.function <- read.csv('Covariates/KCNQ1_functional_data_trimmed_v2.csv', header=TRUE, stringsAsFactors=FALSE)

# merge with in silico list

q1.covariates <- merge(in.silico, q1.function, all=TRUE)

# May optionally take only SNV data using REVEL subset - here we keep all 

save(cohort.data, q1.covariates, lit.cohort.data, lit.nonoverlap.data, mut_type, file="KCNQ1_prepared_data")

```


